Equalizer 2.0\\Convergence of a Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Resolve some fixmes
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@41 db7c848b-7755-4df2-b763-8796797afa52

The concept of transparent OpenGL interception popularized by WireGL and
Chromium~\cite{HHNFAKK:02} has received little attention since 2009. While some
commercial implementations such as TechViz and MechDyne Conduit continue to
exist, on the research side only ClusterGL~\cite{NHM:11} has been
presented. ClusterGL employs the same approach as Chromium, but delivers a
significantly faster implementation of transparent OpenGL interception and
distribution for parallel rendering. CGLX~\cite{DK:11} tries to bring parallel
execution transparently to OpenGL applications, by emulating the GLUT API and
intercepting certain OpenGL calls. In contrast to frameworks like Chromium, CGLX
follows the distributed application approach, which works transparently for
trivial applications, but quickly requires the application developer to address
the complexities of a distributed application when mutable application states
need to be synchronized across processes.\FIXME{do we need to state how Equalizer
is different from the new ClusterGL and CGLX?}

On the other hand, software for driving and interacting with tiled display walls
has received significant attention, including Sage~\cite{Sage} and Sage~2~\cite{Sage2}
in particular. Sage was built entirely around the concept
of a shared framebuffer where all content windows are separate applications
using pixel streaming but is no longer actively supported. Sage 2 is a complete,
browser-centric reimplementation where each application is a web application
distributed across browser instances. DisplayCluster~\cite{DC}, and its
continuation Tide~\cite{tide}, also implement the shared framebuffer concept of
Sage, but provide a few native content applications integrated into the display
servers.\FIXME{also here, maybe a differentiation to Sage2 and DC is needed,
showing the advantages of EQ}

@@

{\em Equalizers} are an addition to compound trees. They modify parameters of their
respective subtree at runtime to dynamically optimize one aspect of the decomposition. Due
to their nature, they are transparent to application developers, but might have
application-accessible parameters to tune their behavior.
\FIXME{ok, this is a bit tricky to have Equalizer as a name for the framework as well as for this component, can't we change this?}

>>>>

The concept of transparent OpenGL interception popularized by WireGL and
Chromium~\cite{HHNFAKK:02} has received little attention since 2009. While some
commercial implementations such as TechViz and MechDyne Conduit continue to
exist, on the research side only ClusterGL~\cite{NHM:11} has been presented.
ClusterGL employs the same approach as Chromium, but delivers a significantly
faster implementation of transparent OpenGL interception and distribution for
parallel rendering. CGLX~\cite{DK:11} tries to bring parallel execution
transparently to OpenGL applications, by emulating the GLUT API and intercepting
certain OpenGL calls. In contrast to frameworks like Chromium and ClusterGL
which distribute OpenGL calls, CGLX follows the distributed application
approach. This works transparently for trivial applications, but quickly
requires the application developer to address the complexities of a distributed
application, when mutable application state needs to be synchronized across
processes. For realistic applications, writing parallel applications remains the
only viable approach for scalable parallel rendering, as shown by the success of
Paraview, Visit and Equalizer-based applications.

On the other hand, software for driving and interacting with tiled display walls
has received significant attention, including Sage~\cite{Sage} and
Sage~2~\cite{Sage2} in particular. Sage was built entirely around the concept of
a shared framebuffer where all content windows are separate applications using
pixel streaming but is no longer actively supported. Sage 2 is a complete,
browser-centric reimplementation where each application is a web application
distributed across browser instances. DisplayCluster~\cite{DC}, and its
continuation Tide~\cite{tide}, also implement the shared framebuffer concept of
Sage, but provide a few native content applications integrated into the display
servers. These solutions implement a scalable display environment and are a
target display platform for scalable 3D graphics applications.

@@

{\em Equalizers} are an addition to compound trees. They modify parameters of
their respective subtree at runtime to dynamically optimize the resource usage,
by each tuning one aspect of the decomposition. Due to their nature, they are
transparent to application developers, but might have application-accessible
parameters to tune their behavior. Resource equalization is the critical
component for scalable parallel rendering, and therefore the eponym for the
Equalizer project name. \FIXME{ok, this is a bit tricky to have Equalizer as a
name for the framework as well as for this component, can't we change this?
eile: We can't change the product name, and changing the component name will be
hard. In fact, this is an artefact of the architecture: The runtime balancing of
resources is the core contribution of Equalizer, and therefore the equalizers
performing this task are the name-givers for the product. Modified paragraph
above for clarity.}
